Software Citation Principles

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex

\textbf{Our response:}
This has been fixed.

\item \textbf{Reviewer comment:}
\emph{Lines 241--242: ``Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.'' I found this non sequitur (or am not seeing the logical connection). I can imagine reasons to cite software that does not form part of the provenance chain of results presented (e.g. alternatives considered).}

\textbf{Our response:}

We agree with the reviewer.  We had been thinking of citation only in the context of software that had been used, not the other uses of citation.  We have changed the text to acknowledge these other uses of citation.

@@

\item \textbf{Reviewer comment:}
\emph{Lines 245--247: ``In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification. Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).'' This statement feels to me to be at odds with table 2, which presents ``Use cases and basic metadata requirements for software citation'', which includes ``reproducibility'' as a use-case. This leaves me wondering if the use cases in table 2 are going beyond what is strictly needed for the purposes of software citation.}

\textbf{Our response:}
We have rephrased this statement to clarify our meaning.

The requirements given in the use cases table only indicate the metadata needed for citation of software for each use case, and not necessarily all the requirements needed to actually complete each use case. For example, use case 7 ``Benchmark software'' does not include requirements for performing actual benchmarking, including hardware, performance monitoring software, performance modeling knowledge, etc.
\niemnote{how does this sound, in combination with the suggested rephrasing?}
\katznote{I've changed the text a little}

### 2016-07-21 sc-principles/software-citation-principles.tex
some wording edits in section 3 (software vs. data)

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
The adaptations were made because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. While software can be considered a type of data, the converse is not
generally true.  In addition:

\begin{itemize}
\item Software can be used to express or explain concepts, unlike data.
\item Software is updated more frequently that papers or data\footnote{Datasets can of course be frequently updated, but each update likely represents the findings of a new or repeated study.}.
\item Software is executable, unlike papers or data.
\item Software suffers from a different type of bit rot than data --- software must be constantly maintained so that it continues to function as both the hardware and software environments that it is used on changes, as bugs are found and fixed, and as user requirements demand new features and capabilities.
\item Software is frequently built to use other software, leading to complex dependencies, and these dependent software packages also are frequently changing.
\item Software is generally smaller than data, so a number of the storage and preservation constraints on data don't apply to software.
\item Software teams can be very large, and very multidisciplinary, and there can be many varied roles that are overlapping and sometimes ambiguous with respect to contributions, and which contributions rise to a level equivalent to paper authorship.
\item The lifetime of software is generally not as long as that of data.
\end{itemize}

The principles in this document should guide further development of software citation
mechanisms and systems, and the reader should be able to look at any particular example
of software citation and see if it meets the principles.
Please note that while we strive to offer practical guidelines that acknowledge the current incentive system of academic citation, a more modern system of assigning credit is sorely needed.
It is not that academic software needs a separate system from academic papers, but that the need for credit for application software underscores the need to overhaul the system of credit for all research products.

>>>>

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
We made the adaptations because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. Software can be considered a type of data, but the converse is not
generally true. In addition:

\begin{itemize}
\item Software can be used to express or explain concepts, unlike data.
\item Software is updated more frequently that papers or data\footnote{Datasets
can of course be frequently updated, but each update likely represents the
findings of a new or repeated study.}.
\item Software is executable, unlike papers or data.
\item Software suffers from a different type of bit rot than data---software
must be constantly maintained so that it continues to function as both the
hardware and software environments that it is used on changes, as developers
find and fix bugs, and as user requirements demand new features and capabilities.
\item Software is frequently built to use other software, leading to complex
dependencies, and these dependent software packages also frequently change.
\item Software is generally smaller than data, so a number of the storage and
preservation constraints on data don't apply to software.
\item Software teams can be large and multidisciplinary, and with varied roles
that overlap and sometimes make contributions ambiguous---some of which
can rise to a level equivalent to paper authorship.
\item The lifetime of software is generally not as long as that of data.
\end{itemize}

The principles in this document should guide further development of software citation
mechanisms and systems, and the reader should be able to look at any particular example
of software citation and see if it meets the principles.
While we strive to offer practical guidelines that acknowledge the current
incentive system of academic citation, a more modern system of assigning credit
is sorely needed. It is not that academic software needs a separate credit
system from that of academic papers, but that the need for credit for
research software underscores the need to overhaul the system of credit for
all research products.
